### 2016-06-27 Projects/DelCo_2012/Report/DelCo_2012.tex
update report and plotting script

\vspace*{2\baselineskip}
\textit{$^+$\small Currently with JENSEN HUGHES}
\\
\textit{$^*$\small Currently with UL FSRI}
}

@@

\begin{tabbing}
\hspace{1.5in} \= \\
CAF \> Compressed Air Foam \\
ESTC \> Delaware County Emergency Services Training Center \\
MCFRS \> Montgomery County (MD) Fire and Rescue Service \\
NIST \> National Institute of Standards and Technology \\
TFT \> Task Force Tips \\
\end{tabbing}

@@

A number of fire departments throughout the US have acquired CAFS units for structural firefighting, but many have limited or stopped the use of these CAF for structural firefighting due to operational challenges or due to lack of understanding related to fireground safety issues. This project is expected to produce a better understanding of the effectiveness and safety implications of CAF for structural firefighting through the various tasks associated with this project, including: 1) A literature review of available published information on the use of CAFS for structural firefighting~\cite{Mitchell:1}; 2) A workshop that brings interested and experienced parties together to discuss safety implications associated with this project and to provide guidance on the proposed experimental research plan~\cite{Grant:2011}; 3) The performance and analysis of a number of well-instrumented large-scale enclosure fire tests using both CAFS and plain water to evaluate the effectiveness of CAFS for interior structural firefighting; and 4) The performance and analysis of a number of fire ground evolutions to evaluate differences between CAFS and plain water hose lines related to issues such as effective hose stream throw and distribution, nozzle reaction force, the force to kink the hose, potential for slipping on a variety of surfaces and flow separation~\cite{Carracino:2013,Dicus:2013,LaPolla:2012}. The research findings are provide information with the potential to improve the safe and effective use of CAFS for structural firefighting.

@@

The key measures were to understand the cooling and suppression capabilities of CAF compared to water. The experimental designs were developed based on a project planning workshop which included the discussion of incidents where firefighters had concerns or experienced minor burns while using CAFS during interior attack operations in residential occupancies~\cite{Grant:2011}. Anecdotally, other fire departments or firefighters had similar incidents or impressions of CAF.

Compressed air foam (CAF) generated with a solution of water and a Class A foam concentrate has been in use by wildland fire agencies since the late 1970's. The wildland fire community found that the Class A foam reduced the surface tension of water, which improved water's ability to penetrate and soak into wildland fuels. A CAFS could turn a Class A foam solution into a blanket of bubbles that would allow the wildland firefighters to coat the fuels. The additional ``soak'' time provided by the foam layer allows the water to penetrate the hot surface of a burning fuel, and also provides additional time for water to penetrate into fuels being pre-treated in an effort to prevent ignition. An added feature of the blanket of bubbles is that firefighters could see where they had applied the agent so that they could make a more effective application. In the wildland, water for firefighting is typically a precious resource, is in limited supply, and has to be carried to where it is needed. CAF increases the effectiveness of water, and may reduce the quantity required to suppress fires and prevent the ignition of adjacent fuels. Last but not least the compressed air adds energy to the water stream providing additional reach and volume to a hose stream with a relatively low mass flow rate of water. These CAFS advantages combine to provide wildland firefighters with a tool that could be used to pre-treat homes and structures to protect them from an approaching fire, or could be used for an exterior (indirect) fire suppression attack on a structure.

In the decades that followed the development of CAFS, a number of urban/suburban fire departments across the country began to integrate CAFS into their apparatus. For example, the Los Angeles County Fire Department incorporated CAFS into their apparatus for wildland urban interface fires, but also examined using them for structural fire fighting. Departments across the country began to experiment with CAFS systems for use in structural firefighting. In most documented cases, the findings about the effectiveness of the CAF were favorable, yet after decades of fire department trials, CAF has still not been significantly incorporated into structural firefighting tactics. This is illustrated by the example of one community, which after putting a fleet of CAFS equipped fire engines into service, had issued an order that the CAFS should not be used after incidents occurred where firefighters were burned while performing interior suppression operations with CAF inside a structure.

There has never been a comprehensive scientific study on the use of CAF for structural fire fighting.  Fire departments have conducted pilot programs to examine implementation and effectiveness of CAF, foam suppliers and equipment manufacturers have conducted minimally instrumented experiments, typically in acquired structures, and test laboratories and universities have examined small sections of the CAF puzzle, but none of these efforts has resulted in the body of knowledge needed by the fire service~\cite{Madrzykowski:3,USFA:CAFS,Colletti:1,Boyd:1,Tafreshi:1,Madrzykowski:4,Colletti:2,Tafreshi:2,Lorh:2002,Colletti:3,Kim:2012}.

@@

Montgomery County (MD) Fire and Rescue Service (MCFRS) is one of the larger fire departments in the State of Maryland and was a committed user of CAFS with 35 CAFS front line engines. MCFRS had experienced two incidents where firefighters received burns in residential structure fires whiles CAF was being used to fight the fire during an interior attack. Concerns were raised within the department regarding the ability of CAF to cool fire gases and absorb heat as well as plain water. As a result, on March 4, 2010, the Fire Chief of (MCFRS) issued a General Order prohibiting the use of CAFS during offensive interior attacks. In this same order, the Chief directed the Division Chief of Operations to continue to evaluate the use of CAFS during structure fires and to report back his findings and recommendations. The catalyst for this study was as a result of this order.

@@

All of the experiments were conducted at full scale in compartments and structures that were representitive room sizes in a residential occupancy.  Technical panel members whose departments rountinely use CAF for structural firefighting were invited and present at many of the fire suppression experiments.  The series of experiments, planned during the workshop, for examining spray density measures, gas layer cooling, and fire suppression were conducted in the fall of 2012.  The data from these experiments were presented to the technical panel member.   The fire suppression experiments were conducted with a fuel package that included upholstered furniture.  After a review of the data, a decision was made to conduct an additional series of suppression experiments with a fuel package was composed mainly of wood based materials, as a way to increase the challenge of the suppression experiment.  After the wood based fuel package experiments were completed, it was decided in discussions with members of the technical panel that other experimental conditions needed to be examined: room size, additional ventilation, and a manually operated handline.  The issue of a larger fire compartment had also been raised durning the workshop discussion, but the desire for replicate experiments and financial limitations removed the large floor plan experiments from the original plan.  In order to accomplish the last series of experiments, additional funding was obtained from NIST and the experiments were conducted in 2015.

The sections below describe the types of experiments, the experimental facilities, instrumentation, and fuel loads.  Project partner, Montgomery County (MD) Fire and Rescue Service (MCFRS) provided the CAFS equipped engines and the staffing needed to provide both the water and CAF for the experiments. These experiments were conducted at the Delaware County Emergency Services Training Center (ESTC) in Sharon Hill, PA.

@@

The fire suppression series included 18 tests in two single-story experimental structures and one two-story structure. Fourteen tests were conducted in the single story structures (7 tests in each structure) using both water and CAFs application as a suppression agent. The nozzle patterns were varied from straight stream to narrow fog. Two of the tests examined suppression tactics for attic fires in residential dwellings, one in each structure. For the attic tests, a smooth bore nozzle was used. The 4 two-story tests used both water and CAF from a straight-stream nozzle.

@@

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Gas Cooling Experiments}\label{tab:spray_density_tests}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    &  Duration (s)  & Agent  &  Pattern            & Nozzle Position  \\
\midrule
GC1        &  15            & Water  &  Solid Stream       & Back             \\
GC2        &  23            & Water  &  Solid Stream       & Back             \\
GC3        &  15            & Water  &  Solid Stream       & Mid              \\
GC4        &  15            & Water  &  Solid Stream       & Mid              \\
GC5        &  20            & Water  &  Solid Stream       & Mid              \\
GC6        &  21            & Water  &  30$^{\circ}$ Fog   & Mid              \\
GC7        &  20            & Water  &  30$^{\circ}$ Fog   & Mid              \\
GC8        &  22            & Water  &  30$^{\circ}$ Fog   & Back             \\
GC9        &  24            & Water  &  30$^{\circ}$ Fog   & Back             \\
GC10       &  20            & Water  &  Solid 7/8~in Slug  & Back             \\
GC11       &  22            & Water  &  Solid 7/8~in Slug  & Back             \\
GC12       &  15            & CAF    &  Smooth Bore 7/8~in & Back             \\
GC13       &  10            & CAF    &  Smooth Bore 7/8~in & Mid              \\
GC14       &  15            & CAF    &  Smooth Bore 7/8~in & Back             \\
GC15       &  20            & CAF    &  Smooth Bore 7/8~in & Mid              \\
GC16       &  15            & CAF    &  30$^{\circ}$ Fog   & Mid              \\
GC17       &  20            & CAF    &  Solid Stream       & Mid             \\
GC18       &  10            & CAF    &  Solid Stream       & Mid             \\
GC19       &  15            & CAF    &  30$^{\circ}$ Fog   & Back             \\
GC20       &  15            & CAF    &  30$^{\circ}$ Fog   & Back             \\
GC21       &  15            & CAF    &  Solid Stream       & Back             \\

@@

\begin{figure}[ht!]
	\includegraphics[width=\columnwidth]{../Figures/Gas_Cooling/GCSeries4_TC_A3}
	\caption{Time series of several gas cooling events in a continuous test}
	\label{fig:gas_cooling_exp4}
\end{figure}

@@

\begin{figure}[ht!]
	\includegraphics[width=.45\columnwidth]{../Figures/Gas_Cooling/GCSeries12_TC_A1}
	\includegraphics[width=.45\columnwidth]{../Figures/Gas_Cooling/GCSeries12_TC_A2}
	\includegraphics[width=.45\columnwidth]{../Figures/Gas_Cooling/GCSeries12_TC_A3}
	\caption{Time series of several gas cooling events in a continuous test}
	\label{fig:gas_cooling_sub1}
\end{figure}

\begin{figure}[ht!]
	\includegraphics[width=.45\columnwidth]{../Figures/Gas_Cooling/GCSeries52_TC_A1}
	\includegraphics[width=.45\columnwidth]{../Figures/Gas_Cooling/GCSeries52_TC_A2}
	\includegraphics[width=.45\columnwidth]{../Figures/Gas_Cooling/GCSeries52_TC_A3}
	\caption{Time series of several gas cooling events in a continuous test}
	\label{fig:gas_cooling_sub5}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_all_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for all replicate tests}
	\label{fig:combined_all}
\end{figure}

The circles in Fig.~\ref{fig:combined_all} are the maximum temperature differences for each measurement point for all the replicate cooling events. If water and CAFs have the exact same temperature differences, all of the points would fall along solid black line in the figure. Circles that vary from that line represent measurements where either water (above) or CAFS (below) cooled more signifcantly than the other agent. The gray area surrounding the solid black line, represents measurement uncertainty (\ref{subsec:Uncertainty}). For data that fall within this area, there is an inability to determine which agent was more or less effective than the other. The red dashed line is the line the describes the data by minimizing the differences between the two suppression agents through a linear approximation of the data. Since this line falls within the gray area there is an inability to distinguish, with statistical confidence, a difference between water or CAFs to cool gases by combining all of the replicate tests.

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/7_8_Solid_mid_A3_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for 7/8 Solid Stream}
	\label{fig:CAFS_Water_Comp_7_8}
\end{figure}

The second water and CAFs comparison was to examine 5 replicate tests for a straight stream in the mid-position, Fig.~\ref{fig:CAFS_Water_Comp_SS_mid}. The same analysis was performed for this configuration. These results also illustrate the inability to discern which suppression agent is more or less effective at gas cooling.

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/SS_mid_A3_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for Straight Stream Mid Position}
	\label{fig:CAFS_Water_Comp_SS_mid}
\end{figure}

@@

\begin{table}[!ht]
\centering
\caption{Fire Suppression Test Descriptions}\label{tab:Test_Descriptions}
\begin{tabular}{clllll}
\toprule[1.5pt]
Test $\#$  & Name	& Stream			& Flow Rate		& Fuel           & Ventilation     \\
\midrule
 1 & CAF~1    &  30$^{\circ}$ Fog  	&  120~gpm/60~cfm   & Furniture      & Door            \\
 2 & Water~1  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door            \\
 3 & CAF~2 	  &  7/8'' Smooth Bore  	&  120~gpm/60~cfm   & Furniture      & Door            \\
 4 & Water~2  &  7/8'' Smooth Bore  	&  120~gpm    		& Furniture      & Door            \\
 5 & Water~3  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 6 & CAF~3    &  30$^{\circ}$ Fog  	&  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 7 & Water~4  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 8 & CAF~4    &  30$^{\circ}$ Fog  	&  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 9 & Attic~1  &  Straight Stream  	&  120~gpm    		& Furniture      & Door            \\
10 & Attic~2  &  Straight Stream  	&  120~gpm/60~cfm   & Furniture      & Door            \\
11 & Water~5  &  30$^{\circ}$ Fog  	&  120~gpm   		& Wood           & Door + Window   \\
12 & CAF~5    &  30$^{\circ}$ Fog  	&  120~gpm/60~cfm   & Wood           & Door + Window   \\
13 & Water~6  &  30$^{\circ}$ Fog  	&  120~gpm    		& Wood           & Door + Window   \\
14 & CAF~6    &  30$^{\circ}$ Fog  	&  120~gpm/60~cfm   & Wood           & Door + Window   \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

Table~\ref{tab:Test_Results} provides the impact of the two nozzle applications for the water and CAFs tests. The table provides change in temperature 0.3~m (1~ft) below the ceiling and 0.3~m (1~ft) above the floor as well as change in heat flux 0.15~m (6~in) above the floor at the measurement location closest to the fire. These values reported in the table are immediately before and after the 15~s application period to assess the application agent, stream, and location on interior conditions.

>>>>

\vspace*{2\baselineskip}
{\em $^+$\small Currently with JENSEN HUGHES}
\\
{\em $^*$\small Currently with UL FSRI}
}

@@

\begin{tabbing}
\hspace{1.5in} \= \\
CAFS \> Compressed Air Foam System \\
ESTC \> Delaware County Emergency Services Training Center \\
MCFRS \> Montgomery County (MD) Fire and Rescue Service \\
NIST \> National Institute of Standards and Technology \\
TFT \> Task Force Tips \\
\end{tabbing}

@@

A number of fire departments throughout the US have acquired CAFS units for structural firefighting, but many have limited or stopped the use of these CAFS for structural firefighting due to operational challenges or due to lack of understanding related to fireground safety issues. This project is expected to produce a better understanding of the effectiveness and safety implications of CAFS for structural firefighting through the various tasks associated with this project, including: 1) A literature review of available published information on the use of CAFS for structural firefighting~\cite{Mitchell:1}; 2) A workshop that brings interested and experienced parties together to discuss safety implications associated with this project and to provide guidance on the proposed experimental research plan~\cite{Grant:2011}; 3) The performance and analysis of a number of well-instrumented large-scale enclosure fire tests using both CAFS and plain water to evaluate the effectiveness of CAFS for interior structural firefighting; and 4) The performance and analysis of a number of fire ground evolutions to evaluate differences between CAFS and plain water hose lines related to issues such as effective hose stream throw and distribution, nozzle reaction force, the force to kink the hose, potential for slipping on a variety of surfaces and flow separation~\cite{Carracino:2013,Dicus:2013,LaPolla:2012}. The research findings are provide information with the potential to improve the safe and effective use of CAFS for structural firefighting.

@@

The key measures were to understand the cooling and suppression capabilities of CAFS compared to water. The experimental designs were developed based on a project planning workshop which included the discussion of incidents where firefighters had concerns or experienced minor burns while using CAFS during interior attack operations in residential occupancies~\cite{Grant:2011}. Anecdotally, other fire departments or firefighters had similar incidents or impressions of CAFS.

Compressed air foam (CAF) generated with a solution of water and a Class A foam concentrate has been in use by wildland fire agencies since the late 1970's. The wildland fire community found that the Class A foam reduced the surface tension of water, which improved water's ability to penetrate and soak into wildland fuels. A CAFS could turn a Class A foam solution into a blanket of bubbles that would allow the wildland firefighters to coat the fuels. The additional ``soak'' time provided by the foam layer allows the water to penetrate the hot surface of a burning fuel, and also provides additional time for water to penetrate into fuels being pre-treated in an effort to prevent ignition. An added feature of the blanket of bubbles is that firefighters could see where they had applied the agent so that they could make a more effective application. In the wildland, water for firefighting is typically a precious resource, is in limited supply, and has to be carried to where it is needed. CAFS increases the effectiveness of water, and may reduce the quantity required to suppress fires and prevent the ignition of adjacent fuels. Last but not least the compressed air adds energy to the water stream providing additional reach and volume to a hose stream with a relatively low mass flow rate of water. These CAFS advantages combine to provide wildland firefighters with a tool that could be used to pre-treat homes and structures to protect them from an approaching fire, or could be used for an exterior (indirect) fire suppression attack on a structure.

In the decades that followed the development of CAFS, a number of urban/suburban fire departments across the country began to integrate CAFS into their apparatus. For example, the Los Angeles County Fire Department incorporated CAFS into their apparatus for wildland urban interface fires, but also examined using them for structural fire fighting. Departments across the country began to experiment with CAFS systems for use in structural firefighting. In most documented cases, the findings about the effectiveness of the CAFS were favorable, yet after decades of fire department trials, CAF has still not been significantly incorporated into structural firefighting tactics. This is illustrated by the example of one community, which after putting a fleet of CAFS equipped fire engines into service, had issued an order that the CAFS should not be used after incidents occurred where firefighters were burned while performing interior suppression operations with CAFS inside a structure.

There has never been a comprehensive scientific study on the use of CAFS for structural fire fighting.  Fire departments have conducted pilot programs to examine implementation and effectiveness of CAFS, foam suppliers and equipment manufacturers have conducted minimally instrumented experiments, typically in acquired structures, and test laboratories and universities have examined small sections of the CAFS puzzle, but none of these efforts has resulted in the body of knowledge needed by the fire service~\cite{Madrzykowski:3,USFA:CAFS,Colletti:1,Boyd:1,Tafreshi:1,Madrzykowski:4,Colletti:2,Tafreshi:2,Lorh:2002,Colletti:3,Kim:2012}.

@@

Montgomery County (MD) Fire and Rescue Service (MCFRS) is one of the larger fire departments in the State of Maryland and was a committed user of CAFS with 35 CAFS front line engines. MCFRS had experienced two incidents where firefighters received burns in residential structure fires whiles CAFS was being used to fight the fire during an interior attack. Concerns were raised within the department regarding the ability of CAFS to cool fire gases and absorb heat as well as plain water. As a result, on March 4, 2010, the Fire Chief of (MCFRS) issued a General Order prohibiting the use of CAFS during offensive interior attacks. In this same order, the Chief directed the Division Chief of Operations to continue to evaluate the use of CAFS during structure fires and to report back his findings and recommendations. The catalyst for this study was as a result of this order.

@@

All of the experiments were conducted at full scale in compartments and structures that were representitive room sizes in a residential occupancy.  Technical panel members whose departments rountinely use CAFS for structural firefighting were invited and present at many of the fire suppression experiments.  The series of experiments, planned during the workshop, for examining spray density measures, gas layer cooling, and fire suppression were conducted in the fall of 2012.  The data from these experiments were presented to the technical panel member.   The fire suppression experiments were conducted with a fuel package that included upholstered furniture.  After a review of the data, a decision was made to conduct an additional series of suppression experiments with a fuel package was composed mainly of wood based materials, as a way to increase the challenge of the suppression experiment.  After the wood based fuel package experiments were completed, it was decided in discussions with members of the technical panel that other experimental conditions needed to be examined: room size, additional ventilation, and a manually operated handline.  The issue of a larger fire compartment had also been raised durning the workshop discussion, but the desire for replicate experiments and financial limitations removed the large floor plan experiments from the original plan.  In order to accomplish the last series of experiments, additional funding was obtained from NIST and the experiments were conducted in 2015.

The sections below describe the types of experiments, the experimental facilities, instrumentation, and fuel loads.  Project partner, Montgomery County (MD) Fire and Rescue Service (MCFRS) provided the CAFS equipped engines and the staffing needed to provide both the water and CAFS for the experiments. These experiments were conducted at the Delaware County Emergency Services Training Center (ESTC) in Sharon Hill, PA.

@@

The fire suppression series included 18 tests in two single-story experimental structures and one two-story structure. Fourteen tests were conducted in the single story structures (7 tests in each structure) using both water and CAFS application as a suppression agent. The nozzle patterns were varied from straight stream to narrow fog. Two of the tests examined suppression tactics for attic fires in residential dwellings, one in each structure. For the attic tests, a smooth bore nozzle was used. The 4 two-story tests used both water and CAFS from a straight-stream nozzle.

@@

\begin{table}[!ht]
\footnotesize
\centering
\captionof{table}{Spray Density Test Matrix for Gas Cooling Experiments}\label{tab:spray_density_tests}
\begin{tabular}{lllll}
\toprule[1.5pt]
Test \#    &  Duration (s)  & Agent  &  Pattern            & Nozzle Position  \\
\midrule
GC1        &  15            & Water  &  Solid Stream       & Back             \\
GC2        &  23            & Water  &  Solid Stream       & Back             \\
GC3        &  15            & Water  &  Solid Stream       & Mid              \\
GC4        &  15            & Water  &  Solid Stream       & Mid              \\
GC5        &  20            & Water  &  Solid Stream       & Mid              \\
GC6        &  21            & Water  &  30$^{\circ}$ Fog   & Mid              \\
GC7        &  20            & Water  &  30$^{\circ}$ Fog   & Mid              \\
GC8        &  22            & Water  &  30$^{\circ}$ Fog   & Back             \\
GC9        &  24            & Water  &  30$^{\circ}$ Fog   & Back             \\
GC10       &  20            & Water  &  Solid 7/8~in Slug  & Back             \\
GC11       &  22            & Water  &  Solid 7/8~in Slug  & Back             \\
GC12       &  15            & CAFS    &  Smooth Bore 7/8~in & Back             \\
GC13       &  10            & CAFS    &  Smooth Bore 7/8~in & Mid              \\
GC14       &  15            & CAFS    &  Smooth Bore 7/8~in & Back             \\
GC15       &  20            & CAFS    &  Smooth Bore 7/8~in & Mid              \\
GC16       &  15            & CAFS    &  30$^{\circ}$ Fog   & Mid              \\
GC17       &  20            & CAFS    &  Solid Stream       & Mid             \\
GC18       &  10            & CAFS    &  Solid Stream       & Mid             \\
GC19       &  15            & CAFS    &  30$^{\circ}$ Fog   & Back             \\
GC20       &  15            & CAFS    &  30$^{\circ}$ Fog   & Back             \\
GC21       &  15            & CAFS    &  Solid Stream       & Back             \\

@@

\begin{figure}[ht!]
	\includegraphics[width=\columnwidth]{../Figures/Gas_Cooling/GCSeries4_TC_A3}
	\caption{Time Series of Multiple Gas Cooling Events From a 1 of 9 Tests}
	\label{fig:gas_cooling_exp4}
\end{figure}

@@

\begin{figure}[ht!]
	\includegraphics[width=.5\columnwidth]{../Figures/Gas_Cooling/GCSeries12_TC_A1}
	\includegraphics[width=.5\columnwidth]{../Figures/Gas_Cooling/GCSeries12_TC_A2}
	\includegraphics[width=.5\columnwidth]{../Figures/Gas_Cooling/GCSeries12_TC_A3}
	\caption{Time Series of Thermocouple Data from 3 Arrays in Gas Cooling Room with Water as Suppression Agent}
	\label{fig:gas_cooling_sub1}
\end{figure}

\begin{figure}[ht!]
	\includegraphics[width=.5\columnwidth]{../Figures/Gas_Cooling/GCSeries52_TC_A1}
	\includegraphics[width=.5\columnwidth]{../Figures/Gas_Cooling/GCSeries52_TC_A2}
	\includegraphics[width=.5\columnwidth]{../Figures/Gas_Cooling/GCSeries52_TC_A3}
	\caption{Time Series of Thermocouple Data from 3 Arrays in Gas Cooling Room with CAFS as Suppression Agent}
	\label{fig:gas_cooling_sub5}
\end{figure}

@@

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for All Replicate Tests}
	\label{fig:combined_all}
\end{figure}

The circles in Fig.~\ref{fig:combined_all} are the maximum temperature differences for each measurement point for all the replicate cooling events. If water and CAFS have the exact same temperature differences, all of the points would fall along solid black line in the figure. Circles that vary from that line represent measurements where either water (above) or CAFS (below) cooled more significantly than the other agent. The gray area surrounding the solid black line, represents measurement uncertainty (\ref{subsec:Uncertainty}). For data that fall within this area, there is an inability to determine which agent was more or less effective than the other. The red dashed line is the line the describes the data by minimizing the differences between the two suppression agents through a linear approximation of the data. Since this line falls above the gray area for the comparison, water shows a better ability to cool gas under the same test criteria as CAFS.

The data can be separated to examine the impact of nozzle position (back or middle position) on the ability to cool gases for each of the suppression agents. Figures~\ref{fig:CAFS_Water_mid} and \ref{fig:CAFS_Water_full} show the cooling comparison for 12 replicates in the back position and 3 replicates in the middle position respectively.

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_mid_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for Middle Position}
	\label{fig:CAFS_Water_mid}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/Combined_fullback_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for Back Position}
	\label{fig:CAFS_Water_full}
\end{figure}

For the 12 middle position replicate comparisons (Fig.~\ref{fig:CAFS_Water_mid}), the results (via the dashed red line) point to water having a slight advantage in gas cooling relative to CAFS. At the back position (Fig.~\ref{fig:CAFS_Water_full}), for the three replicate comparisons, the least squared error line falls within the experimental uncertainty. This indicates an inability to distinguish between the two suppression agents with statistical confidence. The data can be further divided to analyze the impact of the hose stream time on comparison between water and CAFS. Figures~\ref{fig:CAFS_Water_78_mid} and \ref{fig:CAFS_Water_SS_mid} show the results for 7 replicate comparisons of a 7/8~in. solid stream and 5 replicate comparisons of a straight stream respectively; both aimed at the middle position.

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/7_8_Solid_mid_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for 7/8~in. Solid Stream Mid Position}
	\label{fig:CAFS_Water_78_mid}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/SS_mid_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for Straight Stream Mid Position}
	\label{fig:CAFS_Water_SS_mid}
\end{figure}

The 7 replicate tests comparing water and CAFS from a 7/8~in. solid stream nozzle aimed at the middle position show a large bias towards water. For the 5 replicate tests with a straight stream nozzle, the line fit through the data falls right at the edge of where there is still uncertainty regarding an assessment of which suppression agent had more cooling.

The back position can also be decomposed as function of stream. Figure~\ref{fig:CAFS_Water_SS_back} shows a single comparison between water and CAFS for a straight stream in the back position while Fig.~\ref{fig:CAFS_Water_Fog_back} shows a 2 replicate comparison between water and CAFS for a 30$^{\circ}$ Fog in the back position.

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/SS_full_back_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for Straight Stream Back Position}
	\label{fig:CAFS_Water_SS_back}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width=.8\columnwidth]{../Figures/Gas_Cooling/30_Fog_full_back_scatter}
	\caption{CAFS vs Water Gas Cooling Comparison for 30$^{\circ}$ Fog Back Position}
	\label{fig:CAFS_Water_Fog_back}
\end{figure}

For the straight stream comparison (Fig.~\ref{fig:CAFS_Water_SS_back}), there is an inability to distinguish between the suppression agents with statistical confidence. The 30$^{\circ}$ Fog replicate comparison (Fig.~\ref{fig:CAFS_Water_Fog_back})

@@

\begin{table}[!ht]
\centering
\caption{Fire Suppression Test Descriptions}\label{tab:Test_Descriptions}
\begin{tabular}{clllll}
\toprule[1.5pt]
Test $\#$  & Name	& Stream			& Flow Rate		& Fuel           & Ventilation     \\
\midrule
 1 & CAFS~1   &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door            \\
 2 & Water~1  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door            \\
 3 & CAFS~2   &  7/8'' Smooth Bore  &  120~gpm/60~cfm   & Furniture      & Door            \\
 4 & Water~2  &  7/8'' Smooth Bore  &  120~gpm    		& Furniture      & Door            \\
 5 & Water~3  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 6 & CAFS~3   &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 7 & Water~4  &  30$^{\circ}$ Fog  	&  120~gpm    		& Furniture      & Door + Window   \\
 8 & CAFS~4   &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Furniture      & Door + Window   \\
 9 & Attic~1  &  Straight Stream  	&  120~gpm    		& Furniture      & Door            \\
10 & Attic~2  &  Straight Stream  	&  120~gpm/60~cfm   & Furniture      & Door            \\
11 & Water~5  &  30$^{\circ}$ Fog  	&  120~gpm   		& Wood           & Door + Window   \\
12 & CAFS~5   &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Wood           & Door + Window   \\
13 & Water~6  &  30$^{\circ}$ Fog  	&  120~gpm    		& Wood           & Door + Window   \\
14 & CAFS~6   &  30$^{\circ}$ Fog   &  120~gpm/60~cfm   & Wood           & Door + Window   \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

Table~\ref{tab:Test_Results} provides the impact of the two nozzle applications for the water and CAFS tests. The table provides change in temperature 0.3~m (1~ft) below the ceiling and 0.3~m (1~ft) above the floor as well as change in heat flux 0.15~m (6~in) above the floor at the measurement location closest to the fire. These values reported in the table are immediately before and after the 15~s application period to assess the application agent, stream, and location on interior conditions.
